Inferring community characteristics in labelled networks

### 2021-05-27 nips/experiments.tex
Yiannis version

\section{Experiments}
\label{sec:experiments}

@@

\begin{itemize}[leftmargin=*]
	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon book sales about U.S. politics, published close to the presidential election in 2004. Two books are connected if they were frequently co-purchased by customers. Vertex features encode the political affiliation of the author (liberal, conservative or neutral).

@@

\begin{figure}[!h]
	\centering
	\begin{subfigure}[t]{0.28\linewidth}
		\centering
		\includegraphics[width=\linewidth]{polbooks-graph.png}
		\caption{Polbooks}
		\label{fig:polbooks-graph}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.28\linewidth}
		\centering
		\includegraphics[width=\linewidth]{school-graph.png}
		\caption{School}
		\label{fig:school-graph}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.28\linewidth}
		\centering
		\includegraphics[width=\linewidth]{fb-graph.png}
		\caption{Facebook egonet}
		\label{fig:fb-graph}
	\end{subfigure}
	\begin{subfigure}[t]{0.10\linewidth}
		\centering
		\includegraphics[width=0.8\linewidth]{10-vertical-legend.png}
		\caption{Legend}
		\label{fig:10-legend}
	\end{subfigure}
	\caption{Networks laid out and coloured according by inferred block memberships $\hat{y}$ for a given experiment iteration. Visualisation performed using \textit{graph-tool} \cite{peixoto_graph-tool_2014}}
	\label{fig:graphs-all}
\end{figure}
\begin{figure}[!h]
	\centering
	\begin{subfigure}[t]{0.32\linewidth}
		\centering
		\vskip 0pt
		\includegraphics[width=\linewidth]{polbooks-null.png}
		\caption{Polbooks}
		\label{fig:polbooks-null}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.32\linewidth}
		\centering
		\vskip 0pt
		\includegraphics[width=\linewidth]{school-null.png}
		\caption{School}
		\label{fig:school-null}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.32\linewidth}
		\centering
		\vskip 0pt
		\includegraphics[width=\linewidth]{fb-null.png}
		\caption{Facebook egonet}
		\label{fig:fb-null}
	\end{subfigure}
	\caption{Reduced dimension feature-to-block generator weight samples}
	\label{fig:null-all}
\end{figure}
\begin{table}[!h]
	\centering
	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ standard deviation)}
	\label{tab:results}
	\resizebox{\textwidth}{!}{%
		\begin{tabular}{c|ccc|c|cc|ccc}
			Dataset  & $B$ & $D$ & $D'$ & $\bar{S}_e$ & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $c^*$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
			Polbooks & 3 & 3 & -- & $2.250 \pm 0.000$ & $0.563 \pm 0.042$ & $0.595 \pm 0.089$ & -- & -- & -- \\
			School & 10 & 13 & 10 & $1.894 \pm 0.004$ & $0.787 \pm 0.127$ & $0.885 \pm 0.129$ & $1.198 \pm 0.249$ & $0.793 \pm 0.132$ & $0.853 \pm 0.132$ \\
			FB egonet & 10  & 480 & 10 & $1.626 \pm 0.003$ & $1.326 \pm 0.043$ & $1.538 \pm 0.069$ & $0.94 \pm 0.019$ & $1.580 \pm 0.150$ & $1.605 \pm 0.106$
		\end{tabular}
	}
\end{table}
%
We require metrics to assess performance. This can be split into two separate components: the microcanonical SBM fit (concerned with the $b$-samples) and the fit of the feature-to-block generator (concerned with the $\theta$-samples). Starting with the SBM, $S(b)$ (equation \ref{eqn:dl-form}) can be interpreted as the description length of the partition imposed by $b$. It is only natural to divide this quantity by the number of entities (nodes and edges) in our graph $N+E$ to allow for rough comparison between graphs. This defines a simple metric to gauge the fit of the SBM: the description length per entity averaged over the $b$-samples (equation \ref{eqn:mean-dl}):
%
\begin{equation}
	\bar{S}_e \coloneqq \frac{1}{(N+E) |\Tcal_b|} \sum_{t\in \Tcal_b} S \left( b^{(t)} \right)
	\label{eqn:mean-dl}
\end{equation}
%
However, to assess the performance of the feature-to-block predictor, we must partition the vertex set $[N]$ into training and test sets. We choose to randomly partition the vertices on each experiment run such that a constant fraction $f$ of the available vertices go to form our training set $\Gcal_0$ and the remainder are held out to form our test set $\Gcal_1$.
The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain. As $|\Gcal_0| \neq |\Gcal_1|$ in general, we cannot use the un-normalised log target $U$ (equation \ref{eqn:U-form}) for comparison as the total cross-entropy loss is scaled by the size of each set but the prior term stays constant. We therefore must use the average cross-entropy loss over each set (equation \ref{eqn:cross-entropy-loss}):
%
\begin{equation}
	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star \left( \theta^{(t)} \right)
	\qquad \textrm{where} \qquad
	\Lcal_\star \left( \theta^{(t)} \right) \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)}
	\label{eqn:cross-entropy-loss}
\end{equation}
%
Where $\star \in \{0, 1\}$ has been introduced to toggle between training and test sets. Table \ref{tab:results} summarises the results for each experiment.\footnote{For a comprehensive list of the hyper-parameters used for each experiment please see appendix \ref{appdx:hyperparams}} We also apply the dimensionality reduction method on the two higher dimensional datasets (the school and FB egonet). For this we leverage equation \ref{eqn:c-star}, to reduce the dimension from $D$ to $D'$ with $k=1$ to yield the maximal cutoff $c^*$. We then retrain the feature-block predictor using just the retained feature set $\Dcal'$ and report the loss over the training and test sets for the reduced classifier -- denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively. These values are also included in table \ref{tab:results}.

Table \ref{tab:results} already highlights some general trends in the results. Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher than the training loss $\bar{\Lcal}_0$. This is expected as our test set is smaller than the training set and so more susceptible to variability in its construction. Indeed, most of the variance in the evaluation of $\bar{\Lcal}_0$ and $\bar{\Lcal}_1$ comes from the random partitioning of the graph into training and test sets. Secondly, it can be seen that the dimensionality reduction procedure brings the training and test losses closer together. This implies that the features we keep are indeed correlated with the underlying graphical partition and that the approach generalises correctly.

The average description length per entity of the graph $\bar{S}_e$ has very low variance implying the detected communities can be found reliably (to within an arbitrary relabelling of blocks). For reference we plot an inferred partition for each of the graphs on figure \ref{fig:graphs-all}. The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at partitioning the graph into densely connected clusters.

@@

We wish to determine whether the author's political affiliation is a good predictor of the overall network structure. We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral). From, figure \ref{fig:polbooks-null}, we see that all 3 blocks have a distinct political affiliation as their largest positive component. This is strong evidence that political affiliation is indeed the axis which best predicts the 3-way natural partition of the graph into blocks. Furthermore, from table \ref{tab:results} we see that the training and test losses are very similar and both are low in magnitude. This provides further evidence to the claim that political affiliation is the best explanatory variable for the overall network structure.

@@

We choose $B=10$ in line with the total number of school-classes. As before, we sample the block-generator parameters $\theta$ and employ the dimensionality reduction technique with standard deviation multiplier $k=1$ to pick out the top $D'=10$ features. We then plot the weights for the surviving features $d \in \Dcal'$ on figure \ref{fig:school-null}. Immediately, we see that only the pupils' class memberships have survived (1A-5B); gender and teacher/student status have been discarded meaning that these are not good predictors of overall macro-structure.

The vast majority of blocks are composed of a single class. However, some blocks have 2 comparably good classes as their predictor. For example, block 2 contains classes 3A and 3B as its 2 best predictors. This suggests that the social divide between classes is less pronounced for pupils in year 3. Conversely, some classes are found to extend over two detected blocks (class 2B spans blocks 6 and 7) but we nonetheless do not have a feature which explains the division. The most surprising block is number 5 - which has comparable weightings for classes 5A and 1B. Perhaps there was a joint event between those two classes on the day the data were collected.

@@

We choose $B=10$ and $D'=10$ for this experiment. The remaining features (figure \ref{fig:fb-null}) are those that best explain the high-level community structure. The majority of the surviving features are education related. Nevertheless, for $D'=10$ we only have good explanations for the makeup of some of the detected blocks; several blocks in figure \ref{fig:fb-null} do not have high-magnitude components for $D'=10$.

When the feature dimension is very large, it becomes increasingly likely that a particular feature may uniquely identify a small set of nodes. If these nodes are all part of the same community then the classifier will overfit for that particular parameter. The regularisation term imposed by the prior goes some way to alleviating this problem. Nevertheless, we see in figure \ref{fig:fb-null} that the feature \verb*|birthday-5| has a very high weight as it relates to block 1 -- but it would be preposterous to conclude that birthdays determine graphical structure. The analyst must remain vigilant of such problems.

>>>>

\section{Experimental results}
\label{sec:experiments}

@@

\begin{itemize}[leftmargin=*]
	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon book sales about U.S. politics, published close to the presidential election in 2004. Two books are connected if they were frequently co-purchased by customers. Vertex features encode the political affiliation of the author (liberal, conservative, or neutral).

@@

\begin{figure}[!h]
	\centering
	\begin{subfigure}[t]{0.28\linewidth}
		\centering
		\includegraphics[width=\linewidth]{polbooks-graph.png}
		\caption{Polbooks}
		\label{fig:polbooks-graph}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.28\linewidth}
		\centering
		\includegraphics[width=\linewidth]{school-graph.png}
		\caption{School}
		\label{fig:school-graph}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.28\linewidth}
		\centering
		\includegraphics[width=\linewidth]{fb-graph.png}
		\caption{Facebook egonet}
		\label{fig:fb-graph}
	\end{subfigure}
	\begin{subfigure}[t]{0.10\linewidth}
		\centering
		\includegraphics[width=0.8\linewidth]{10-vertical-legend.png}
		\caption{Legend}
		\label{fig:10-legend}
	\end{subfigure}
	\caption{Networks laid out and coloured according by inferred block memberships $\hat{y}$ for a given experiment iteration. Visualisation performed using \textit{graph-tool} \cite{peixoto_graph-tool_2014}.}
	\label{fig:graphs-all}
\end{figure}
\begin{figure}[!h]
	\centering
	\begin{subfigure}[t]{0.32\linewidth}
		\centering
		\vskip 0pt
		\includegraphics[width=\linewidth]{polbooks-null.png}
		\caption{Polbooks}
		\label{fig:polbooks-null}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.32\linewidth}
		\centering
		\vskip 0pt
		\includegraphics[width=\linewidth]{school-null.png}
		\caption{School}
		\label{fig:school-null}
	\end{subfigure}
	\hfill
	\begin{subfigure}[t]{0.32\linewidth}
		\centering
		\vskip 0pt
		\includegraphics[width=\linewidth]{fb-null.png}
		\caption{Facebook egonet}
		\label{fig:fb-null}
	\end{subfigure}
	\caption{Reduced dimension feature-to-block generator weight samples.}
	\label{fig:null-all}
\end{figure}
\begin{table}[!h]
	\centering
	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ standard deviation).}
	\label{tab:results}
	\resizebox{\textwidth}{!}{%
		\begin{tabular}{c|ccc|c|cc|ccc}
			Dataset  & $B$ & $D$ & $D'$ & $\bar{S}_e$ & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $c^*$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
			Polbooks & 3 & 3 & -- & $2.250 \pm 0.000$ & $0.563 \pm 0.042$ & $0.595 \pm 0.089$ & -- & -- & -- \\
			School & 10 & 13 & 10 & $1.894 \pm 0.004$ & $0.787 \pm 0.127$ & $0.885 \pm 0.129$ & $1.198 \pm 0.249$ & $0.793 \pm 0.132$ & $0.853 \pm 0.132$ \\
			FB egonet & 10  & 480 & 10 & $1.626 \pm 0.003$ & $1.326 \pm 0.043$ & $1.538 \pm 0.069$ & $0.94 \pm 0.019$ & $1.580 \pm 0.150$ & $1.605 \pm 0.106$
		\end{tabular}
	}
\end{table}
%
We require metrics to assess performance. This can be split into two separate 
components: the microcanonical SBM fit (concerned with the $b$-samples) and 
the fit of the feature-to-block generator (concerned with the $\theta$-samples).

Starting with the SBM, recall that the quantity $S(b)$ in~(\ref{eqn:dl-form}) 
can be interpreted as an ideal ``description length'' of the partition 
imposed by $b$. We define a simple
metric $\bar{S}_e$ to gauge the fit of the SBM,
as the description length per entity 
(i.e., divided by the total number $N+E$ of 
nodes plus edges),
averaged over the $b$-samples:
%
\begin{equation}
	\bar{S}_e \coloneqq \frac{1}{(N+E) |\Tcal_b|} \sum_{t\in \Tcal_b} S \left( b^{(t)} \right).
	\label{eqn:mean-dl}
\end{equation}
%

Next, in order to assess the performance of the feature-to-block predictor, 
we partition the vertex set $[N]$ into training and test sets. We choose to 
randomly partition the vertices on each experiment run,
so that a constant fraction $f$ of the available vertices go to form 
the training set $\Gcal_0$ and the remainder are held out to form the
test set $\Gcal_1$.
The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain. Because $|\Gcal_0| \neq |\Gcal_1|$ in general, we cannot use the un-normalised log-target $U$ from~(\ref{eqn:U-form}) for comparison,
as the total cross-entropy loss scales with the size of each data set but 
the prior term stays constant. We therefore use the average cross-entropy loss 
over each set,
%
\begin{equation}
	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star \left( \theta^{(t)} \right),
	\quad \textrm{where} \quad
	\Lcal_\star \left( \theta^{(t)} \right) \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)},
	\label{eqn:cross-entropy-loss}
\end{equation}
%
and where $\star \in \{0, 1\}$ indicates whehter the training or test
set is being considered.

Table~\ref{tab:results} summarises the results for each experiment;
a complete list of all the hyperparameters used 
is given in Appendix \ref{appdx:hyperparams}. We also apply the 
dimensionality reduction method 
of Section~\ref{sec:dim-reduction}
to the two higher dimensional datasets (the school and FB egonet). 
For this we use equation~(\ref{eqn:c-star}) with $k=1$,
in order to reduce the dimension from 
$D$ to a desired $D'$ with $k=1$ based on the maximal cutoff $c^*$. 
We then retrain the feature-block predictor using only the retained 
feature set $\Dcal'$, and report the log-loss over the training and 
test sets for the reduced classifier -- 
denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively. 
These values are also shown in Table~\ref{tab:results}.

Based on these results, some remarks are in order.
Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher 
than the training loss $\bar{\Lcal}_0$. This is expected,
as the test set is smaller than the training set and hence 
more susceptible to variability in its construction. Indeed, 
most of the variance in the evaluation of $\bar{\Lcal}_0$ and $\bar{\Lcal}_1$ 
comes from the random partitioning of the graph into training and test sets. 
Secondly, it can be seen that the dimensionality reduction procedure 
brings the training and test losses closer together. This implies that 
the features we keep are indeed correlated with the underlying graphical 
partition and that the approach generalises correctly.

The average description length per entity,
$\bar{S}_e$, of the graph, 
has very small variance, suggesting that
the detected communities can be found reliably (to within an arbitrary 
relabelling of blocks). For reference, we plot an inferred partition for each 
of the graphs in Figure~\ref{fig:graphs-all}. The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at partitioning the graph into densely connected clusters.

@@

The goal here is to 
determine whether the authors' political affiliations are a good predictor 
of the overall network structure. We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal. or neutral). From Figure~\ref{fig:polbooks-null} we see that all 3 blocks have a distinct political affiliation as their largest positive component. 
This strongly indicates that political affiliation is indeed the axis 
which best predicts the 3-way natural partition of the graph into blocks. 
Furthermore, in Table~\ref{tab:results} we see that the training and test losses 
are very similar and both are low in magnitude. This provides further evidence 
to the claim that political affiliation is a very appropriate explanatory 
variable for the overall network structure.

@@

We choose the number of communities $B=10$, in line with the total number of 
school classes. As before, we sample the block-generator parameters $\theta$ and 
employ the dimensionality reduction technique 
of Section~\ref{sec:dim-reduction},
with standard deviation multiplier $k=1$ to pick out the top $D'=10$ features. 
We then plot the weights for the resulting features $d \in \Dcal'$ 
in Figure~\ref{fig:school-null}. It is rather obvious
that only the pupils' class memberships (1A-5B) have remained
significant;
gender and teacher/student status have been discarded,
meaning that these are not good predictors of overall macro-structure.

The vast majority of blocks are composed of a single class. 
However, some blocks have two comparably good classes as their predictor. 
For example, block 2 contains classes 3A and 3B as its 2 best predictors. 
This suggests that the social divide between classes is less pronounced 
for pupils in year 3. Conversely, some classes are found to extend over two 
detected blocks (class 2B spans blocks 6 and 7) but we nonetheless do 
not have a feature which explains the division. The most surprising block 
is number 5, which has comparable weightings for classes 5A and 1B. 
Perhaps there was a joint event between those two classes on the day 
the data were collected.

@@

We choose $B=10$ and $D'=10$ for this experiment. The selected features 
(Figure~\ref{fig:fb-null}) are those that best explain the high-level 
community structure. The majority of them are education related. 
Nevertheless, for $D'=10$ we only have good explanations for the makeup 
of some of the detected blocks; several blocks in 
Figure~\ref{fig:fb-null} do not have high-magnitude components for $D'=10$.

When the feature dimension is very large, it becomes increasingly likely that 
a particular feature may uniquely identify a small set of nodes. If these nodes 
are all part of the same community, then the classifier will overfit for that 
particular parameter. The regularisation term imposed by the prior goes some 
way towards alleviating this problem. Nevertheless, we see in 
Figure~\ref{fig:fb-null} that the feature \verb*|birthday-5| has a very 
high weight as it relates to block 1 -- but it is highly unlikely
that birthdays determine graphical structure. 
% The analyst must remain vigilant of such problems.

### 2021-05-27 nips/ffbm.tex
Yiannis version

In this section we propose a novel generative model for labelled networks. We call this the feature-first block model (FFBM) and outline its structure in \ref{fig:ffbm} As before, we let $N$ denote the number of nodes and $B$ the number of blocks in our graph. We define the vector $x_i \in \Xcal^D$ as the feature vector for the $i$'th vertex. $D$ is the number of features. For the datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$. The feature vectors $\{x_i\}_{i=1}^{N}$ may be compactly subsumed into the feature matrix $X \in \Xcal^{N \times D}$.

For the FFBM, we start with the feature matrix X and probabilistically generate a vector of block memberships $b \in [B]^N$. The parameters of this step are encapsulated by $\theta$. Each feature vector $x_i$ is treated independently and used to generate the corresponding block membership $b_i \in [B]$. We choose a single softmax layer to model $p(b_i | x_i, \theta)$. More complex models are possible but then deriving meaning from the inferred parameter distributions is more difficult. Summarising, we write $p(b | X, \theta)$ as follows:
%
\begin{equation}
	p(b| X, \theta) = \prod_{i=1}^{N} p(b_i | x_i, \theta) = \prod_{i=1}^{N} \phi_{b_i} (x_i; \theta)
	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T x_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T x_i\right)}
\end{equation}
%
We deliberately exclude a bias term to ensure that the relationships we model are based on features and not information about the size of each detected block; a more complete discussion on this topic is given in \ref{appdx:dimension}. The parameter vector $\theta$ for this stage contains all the weight vectors $\theta = \{w_k\}_{k=1}^{B}$. Each $w_k$ has dimension $D$. We could instead write the parameters $\theta$ as a $B \times D$ matrix of weights $W$; this form has computational benefits as then $z_i \coloneqq W x_i$, which is the input to the softmax activation function.

@@

Once the block memberships $b$ have been generated, we then draw the graph $A$ from the microcanonical DC-SBM (equation \ref{eqn:A-generation}) with additional parameters encapsulated by $\psi = \{\psi_e, \psi_k\}$.
%
\begin{equation}
	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k)
	\label{eqn:A-generation}
\end{equation}

@@

Before performing any inference, we must specify priors on $\theta$ and $\psi$. For $\theta$ it seems sensible to choose a Gaussian prior, with zero mean and variance matrix $\sigma^2_\theta I$ such that each element of $\theta$ is independent and distributed like $\sim \Gaussian(0, \sigma_\theta^2)$. In vector form, the prior for $\theta$ is therefore:
%
\begin{equation}
	p(\theta) = \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right)
	\label{eqn:theta-prior}
\end{equation}
%
In our model, the block memberships vector $b$ is an intermediate latent variable and so we are not free to choose a prior for it. Nevertheless, as far as inference on the right-hand-side of figure \ref{fig:ffbm}, we regard $p(b | X)$ as a pseudo-prior on $b$. We can show (appendix \ref{appdx:b|x}) that our choice of prior for $p(\theta)$ in equation \ref{eqn:theta-prior} leads to a uniform $p(b | X)$ in equation \ref{eqn:b-pseudo-prior}.
%
\begin{equation}
	p(b | X) = \int p(b | X, \theta) p(\theta) d\theta = B^{-N}
	\label{eqn:b-pseudo-prior}
\end{equation}
%
This is an enormously important simplification as evaluating $p(b | X)$ does not require an expensive Monte-Carlo integration over the $\theta$-domain nor does it require the exact value of $X$. \citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the additional microcanonical SBM parameters $\psi$ which we adopt. Peixoto's idea is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$. For our purposes we must insert a conditioning on $X$, to form our pseudo-prior for $b$ and $\psi$, to give equation \ref{eqn:joint-pseudo-prior}.
%
\begin{equation}
	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b)
	\label{eqn:joint-pseudo-prior}
\end{equation}
%
Where we leverage the fact $(\psi \indep X) | b$. We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$ to complete our model. Please refer to appendix \ref{appdx:sbm} for the exact form of $p(\psi | b)$. All that concerns the main argument is we have a computable form.

>>>>

In this section we propose a novel generative model for labelled networks. We call this the feature-first block model (FFBM).

Let $N$ denote the number of nodes, $B$ the number of blocks in the graph,
and ${\cal X}$ the set of all possible features.
We define the vector $x_i \in \Xcal^D$ as the feature vector of vertex $i$, 
where $D$ is the number of features associated with each vertex.
For the datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$. We write $X$ for the $N\times D$ {\em feature matrix} containing
the feature vectors $\{x_i\}_{i=1}^{N}$ 
as its rows.

For the FFBM, we start with the feature matrix $X$ and generate a random
vector of block memberships $b \in [B]^N$. For each node $i$, the
block membership $b_i\in[B]$ is generated based on the feature
vector $x_i$, independently between nodes. The conditional
distribution of $b_i$ given $x_i$ also depends on a collection
of weight vectors $\theta=\{w_k\}_{k=1}^B$, where each
$w_k$ has dimension $D+1$. Specifically, 
the distribution of $b$ given $X$ and $\theta$ is,
%
\begin{equation}
	p(b| X, \theta) = \prod_{i=1}^{N} p(b_i | x_i, \theta) = \prod_{i=1}^{N} \phi_{b_i} (x_i; \theta)
	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T x_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T x_i\right)}
\end{equation}
%
We deliberately exclude a bias term to ensure that the relationships we model are based on features and not information about the size of each detected block;
a more complete discussion on this topic is given in \ref{appdx:dimension}. 
We will later also find it convenient
to write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has computational benefits as then $z_i \coloneqq W x_i$, which is the input to the softmax activation function.

More complex models based on different choices for the distributions
$\phi_{b_i}$ above are also possible, but then deriving meaning from 
the inferred parameter distributions is more difficult.

@@

Once the block memberships $b$ have been generated, we then draw the 
graph $A$ from the microcanonical DC-SBM with additional parameters 
$\psi = \{\psi_e, \psi_k\}$:
%
\begin{equation}
	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k).
	\label{eqn:A-generation}
\end{equation}

@@

To complete the description of our Bayesian framework,
priors on $\theta$ and $\psi$ must also be specified. 
We place a Gaussian prior on $\theta$ with zero mean
and covariance matrix $\sigma^2_\theta I$, so that
each element of $\theta$ has an independent ${\cal N}(0,\sigma_\theta^2)$
prior, for a hyperparameter $\sigma_\theta^2$:
%
\begin{equation}
	p(\theta) \sim \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right).
	\label{eqn:theta-prior}
\end{equation}
%
Interestingly, this choice of prior leads to a particularly
simple distribution on the block membership vector $b$.
We show in Appendix~\ref{appdx:b|x} that,
after integrating out $\theta$,
$b$ is uniformly distributed:
%
\begin{equation}
	p(b | X) = \int p(b | X, \theta) p(\theta) d\theta = B^{-N}.
	\label{eqn:b-pseudo-prior}
\end{equation}
%
This is an important simplification as evaluating $p(b | X)$ does not require an expensive Monte Carlo integration over $\theta$ nor does it require the exact value of $X$. 
\citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for 
the additional microcanonical SBM parameters $\psi$, which we adopt. 
The idea is to write the joint distribution on $(b, e, k)$ as a product of 
conditionals, $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$. 
In our case, conditioning on $X$ is also necessary, leading to,
%
\begin{equation}
	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b),
	\label{eqn:joint-pseudo-prior}
\end{equation}
%
where we used the fact $\psi$ and $X$ are conditionally 
independent given $b$.  The exact form of the prior
$p(\psi|b)$ is described in Appendix \ref{appdx:prior}.
All that concerns the main argument is that it has
an easily computable form.
